---
title: "Recruitment Simulation"
author: "Xiaojing Zhu"
date: "9/22/2019"
output: html_document
---

## Description of the Problem

In a cluster-randomized trial, the sample size is calculated by the formula $N=n(1+(m-1)\rho)$, where $n$ is the sample size without clustering, $m$ is the cluster size, and $\rho$ is the intra-cluster correlation coefficient (ICC). 

Since cluster sizes, or practice sizes in this case, are usually not constant across medical practices, we need to determine whether the variability of practice sizes is acceptable based on the coefficient variation (CV), $\text{CV}=\frac{\text{SD}}{\text{Mean}}$.

In this problem setting, we plan to enroll 6000 to 6100 patients from 86 medical practices across 10 healthcare system. We consider the following scenarios: the practice sizes either follow a normal distribution with a mean of 70 and varying standard deviations of 5, 10, 15 and 20, or they follow a uniform distribution with a mean of 70, varying minimums of 40, 45 and 50, and varying maximums of 90, 95 and 100. Therefore, a total of 7 scenarios are explored.

To address the problem, the first step is to write a function to output a vector of practice sizes across 86 medical practices. The second step is to write a function to calculate the CV for a vector of practice sizes. The third step is to write another function to create 1000 simulations based on the former two functions. If the $90^{\text{th}}$ percentile of the 1000 CVs are smaller than 0.23, we accept the variability in practice sizes. Otherwise, we reject the variability in practice sizes since the variability is too large.

Finally, 7 tables of output are printed to explain the variability in practice sizes corresponding to the 7 scenarios of interest.

## Step 1. Write a function "PS" to obtain a vector of practice sizes

**Documentation for PS function**

* Function: PS

* Author: Xiaojing Zhu

* Creation Date: Dec 17, 2018 (R version 3.5.1)

* Purpose: This function returns a vector of practice sizes, consisting of the number of participants from each medical practice. This function supports two possible distributions of the practice sizes - either a normal distribution or a uniform distribution. The total sample size, which is the sum of the practice sizes, is no more than 6100 and no less than 6000.

* Required Parameters: 
    - dist = the distribution of practice sizes - a character value of either "Normal" or "Uniform"
  
* Optional Parameters: 
    - no_prac = the total number of medical practices - default is 86
    - mean = the mean of the normal distribution if entered "Normal" - default is 70
    - sd = the standard deviation of the normal distribution if entered "Normal" - default is 5
    - min = the minimum of the uniform distribution if entered "Uniform" - default is 40
    - max = the maximum of the uniform distribution if entered "Uniform" - default is 100

* Output: Return a vector of practice sizes - default length is 86.

* Example: PS(dist = "Normal", mean = 70, sd = 20) returns a vector of practice sizes following a normal distribution with mean of 70 and standard deviation of 20.

```{r}
PS <- function(no_prac = 86, dist, mean = 70, sd = 5, min = 40, max = 100){
  # Check if the distribution is either "Normal" or "Uniform"
  if (! dist %in% c("Normal","Uniform")){ 
    stop("Invalid distribution. Please enter either Normal or Uniform.")
  }
  repeat{
    # Repeat until a total sample size between 6000 and 6100 is found
    # If practice sizes follow a normal distribution:
    if (dist == "Normal") {
      # Round up the number of participants
      ps <- ceiling(rnorm(n = no_prac, mean = mean, sd = sd))
    }
    # If practice sizes follow a uniform distribution:
    else if (dist == "Uniform") {
      # Round up the number of participants
      ps <- ceiling(runif(n = no_prac, min = min, max = max))
    }
    # Make sure the total sample size is no more than 6100 and no less than 6000
    if (sum(ps) >= 6000 & sum(ps) <= 6100) {
      break
    }
  }
  # Return a vector of practice sizes - number of participants across medical practices
  return(ps)
}
```

## Step 2. Write a function "CV" to calculate oefficient of variation for a vector of data

**Documentation for CV function**

* Function: CV

* Author: Xiaojing Zhu

* Creation Date: Dec 17, 2018 (R version 3.5.1)

* Purpose: This function calculates coefficient of variation (CV) of practice sizes after one simulation of 
cluster-randomzied trial. CV is obtained by dividing the standard deviation of practice sizes by the mean of practice sizes.

* Required Parameters: 
    - data: A vector of practice sizes
    
* Output: Return a numeric value of CV

* Example: CV(data = PS(dist = "Uniform", min = 50, max = 90)) returns the CV for a vector of practice sizes following a uniform distribution with minimum of 50 and maximum of 90 if the PS function is previously called.

```{r}
CV <- function(data){
  if(is.vector(data) == FALSE) {
    stop("Please enter a vector for the data parameter")
  }
  cv <- sd(data)/mean(data)
  return(cv)
}
```

## Step 3. Embed "PS" and "CV" functions in the "simulation" function to create 1000 simulations 

**Documentation for simulation function**

* Function: simulation

* Author: Xiaojing Zhu

* Creation Date: Dec 17, 2018  (R version 3.5.1)

* Purpose: This function determines whether the variability in practice sizes is acceptable after 1000 simulations of cluster-randomzied trials. The variability is acceptable if 90% of the simulations have coefficient of variation (CV) smaller than 0.23.

* Required Parameters: 
    - dist = the distribution of practice sizes - a character value of either "Normal" or "Uniform"
  
* Optional Parameters: 
    - seed = the seed to initiate the random number generator and make reproducible results - default is 2018
    - nsim = the number of simulations - default is 1000
    - no_prac = the total number of medical practices - default is 86
    - mean = the mean of the normal distribution if entered "Normal" - default is 70
    - sd = the standard deviation of the normal distribution if entered "Normal" - default is 5
    - min = the minimum of the uniform distribution if entered "Uniform" - default is 40
    - max = the maximum of the uniform distribution if entered "Uniform" - default is 100
      
* Other Functions: Require the PS and CV functions to be called prior to the simulation function

* Library: Require the knitr library to output the table using the kable function.

* Output: Return a table containing the type of the scenario, average CV, 90% cutoff, and acceptability decision based on default of 1000 simulations

* Example: simulation(dist = "Uniform", min = 40, max = 100) returns the table of output when the practice sizes follow a uniform distribution, with a minimum of 40 and a maximum of 100, after 1000 simulations using a seed of 2018.

```{r}
simulation <- function(seed = 2018, nsim = 1000, no_prac = 86, 
                       dist, mean = 70, sd = 5, min = 40, max = 100){
  # Check if "knitr" package is installed in order to use the kable function
  if (!require("knitr")) {
    install.packages("knitr") 
    }
  # Load the package
  require(knitr)
  # Check if the distribution is either "Normal" or "Uniform"
  if (! dist %in% c("Normal","Uniform")){ 
    stop("Invalid distribution. Please enter either Normal or Uniform.")
    }
  # Set the seed - 2018 by default - so that the result is reproducible
  set.seed(seed)
  # Replicate the CV function by nsim times using data from the PS function
  results <- replicate(nsim, 
                       CV(PS(no_prac = no_prac, dist = dist, 
                             mean = mean, sd = sd, min = min, max = max)))
  # If 90th percentile of CVs after 1000 simulations is smaller than 0.23, 
  # we accept this variability in practice sizes
  results <- sort(results)
  if (quantile(results, 0.9) < 0.23){ 
    decision <- "Accept"
    }
  # Otherwise, we reject the variability in practice sizes
  else { 
    decision <- "Reject"
    }
  # Create a table to output scenario, average CV, 90% cutoff, 
  # and acceptability decision based on default of 1000 simulations
  # If the scenario is a normal distribution, print the following table:
  if (dist == "Normal") {
    df <- data.frame(paste0(dist, " distribution with mean ", mean, " and SD ", sd),
                     mean(results), 
                     unname(quantile(results, 0.9)), 
                     decision)
  }
  # If the scenario is a uniform distribution, print the following table:
  else if (dist == "Uniform") {
    df <- data.frame(paste0(dist, " distribution with min ", min, " and max ", max),
                     mean(results), 
                     unname(quantile(results, 0.9)), 
                     decision)
  }
  names(df) <- c("Scenario", "Average CV", "90% Cutoff", "Decision")
  kable(df, digits = 3)
}
```

## Results: 

### Scenario 1
```{r message=FALSE}
simulation(dist = "Normal", mean = 70, sd = 5)
```

### Scenario 2
```{r message=FALSE}
simulation(dist = "Normal", mean = 70, sd = 10)
```

### Scenario 3
```{r message=FALSE}
simulation(dist = "Normal", mean = 70, sd = 15)
```

### Scenario 4
```{r message=FALSE}
simulation(dist = "Normal", mean = 70, sd = 20)
```

### Scenario 5
```{r message=FALSE}
simulation(dist = "Uniform", min = 40, max = 100)
```

### Scenario 6
```{r message=FALSE}
simulation(dist = "Uniform", min = 45, max = 95)
```

### Scenario 7
```{r message=FALSE}
simulation(dist = "Uniform", min = 50, max = 90)
```
